---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      fig.path = "README-")

library(tibble)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

# Ergodicity Economics

## Replicating Example Gamble
First objective was to replicate Fig. 2 of this [paper](https://rdcu.be/cS2t3).

## Code Optimization
In order to try to optimise code for a larger tibble, I have momentarily divided the code in 3 different phases:
- Tibble Generation
- Transformation
- Plotting

The code below is taking the same input given to the paper but instead of 150 players with 1k.

### Tibble Generation
The bottleneck here is replicate, which is nesting sapply, that is nesting lapply. So with a large number of players a lot of vectors are created through lapply. 

```{r, warning = FALSE}
assign("EV", ((0.6 * 0.5) + (1.5 * 0.5)), envir = .GlobalEnv)
df <- as_tibble(replicate(150,
                          cumprod(sample(
                            c(0.6, 1.5),
                            size = 1000,
                            replace = TRUE
                          ))))
```

### Transformation
The bottleneck is on making the tibble longer in order to lett ggplot properly work.

```{r, results = "hide"}
df <- df |>
  mutate(ensemble_avg = rowMeans(df),
         ensemble_median = apply(df, 1, median)) |>
  rowid_to_column("rounds") |>
  mutate(expected_value = EV ^ rounds) |>
  pivot_longer(cols = !1, names_to = "individual")
```

### Plotting
Plotting is the biggest bottleneck.

```{r}
focus <- c("ensemble_avg", "expected_value", "ensemble_median")

ggplot() +
  geom_line(
    subset(df, !individual %in% focus),
    mapping = aes(x = rounds, y = value, group = individual),
    color = "grey",
    size = 0.5,
    alpha = 0.2
  ) +
  geom_line(
    subset(df, individual %in% focus),
    mapping = aes(x = rounds, y = value, color = individual),
    size = 1
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x)
      10 ^ x),
    labels = trans_format("log10", math_format(10 ^ .x))
  ) +
  labs(color = "") +
  theme_bw() +
  theme(legend.position = "bottom")
```









